---
title: "Project Proposal"
author: "Yoni Ackerman"
output: pdf_document
---

```{r, echo = FALSE, warn = FALSE}
library(dplyr)
library(plyr)
library(ggplot2)
library(reshape2)
library(raster)

col_diff <- function(df, col){
    df[, paste0(col, "_diff")] <- c(NA, diff(df[,col]))
    df
}

load("all_data.rda")
load("district_boundaries/dData.rda")
```

# Data

I will be analyzing a political dataset assembled from multiple governmental and
academic sources. The data are as follows:

## State Congress Party Membership
### source: [link](https://dataverse.harvard.edu/)

These are yearly data collected from each state in the union between 1939 and 2015,
inclusive. The available data are organized by year, state, and election year, and
present 96 other cofactors related to that state's government, house and senate,
in that year. Of particular interest to me are variables counting the number
of Republican, Democrat, Independent, Vacant and Total seats in each chamber of
congress. To get an idea of what these data look like, we can examine just
the state senates and how "right" or "left" of center they are:

```{r, echo = FALSE, fig.cap = "Yearly State Congress Party Distribution"}
pData <- pData %>%
    dplyr::mutate(
        state = as.character(state),
        ofCenter = sen_rep_in_sess - sen_dem_in_sess 
        ) %>%
    dplyr::filter(state != "Nebraska")

pData <- ddply(pData, .(state), function(df){
    df <- df %>% dplyr::arrange(year)
    df <- col_diff(df, "ofCenter")
})

ggplot(pData) +
    geom_line(aes(x = year, y = ofCenter, group = state, color = state)) +
    theme(legend.position="none")
```

## District Shapefile Data
### source: [link](http://cdmaps.polisci.ucla.edu/)

These are yearly data that describe the geographical shape and extent of
every State Congressional District in the union, as well as the congress
at the time the data were collected. My goal is to intersect each district's
polygon with it's future polygon's in the future to measure (i) the extent
of the spatial change from year to year and (ii) the cumuluative change over
the course of all measured years. I do not yet have the intersection script
working (I'm not very familiar with R's functionality for spatial data), but
I can present a substitute: yearly area measurements for each district.

```{r, echo = FALSE, fig.cap = "Changes in congression district area for the state of Iowa, by Congress"}

tmp <- dData %>%
    dplyr::filter(state == "Iowa")

ggplot(tmp, aes(x = strt_cng, y = area_sqkm, group = district, color = district)) +
	    geom_line() +
	    theme(legend.position="none")	
```

## District Demographic Data
### source: [link](https://sites.google.com/a/colorado.edu/adler-scott/data)

These data provide yearly demographic measurements for all state congressional
districts in the union, such as total population, number employed, and presence
of a nuclear plant. I am mainly interested in the total population for each
district:

```{r, echo = FALSE, fig.cap = "Changes In Total Population for Congressional Districts in Iowa, by Congress"}

tmp <- dPopData %>%
    dplyr::filter(STATE == "IA")

ggplot(tmp, aes(x = CONGRESS, y = POPULATN, group = CD, color = CD)) +
	    geom_line() +
	    theme(legend.position="none")	   
```

# Objectives

I will analyze how decisions to change or maintain congressional districts
have changed over the past century. Redistricting occurs every 10 years and the lines
are drawn by members of the state congress, thus trends in state congressional
affiliation, as well as prior knowledge of each district's demographics, may
contain both predictive, and explanatory results.

# Anticipated Analysis

So far,
I anticipate using autoregressive models to describe changes in state congress, then
using the residuals from those models. I'd like to then perform frequency analysis
on the 10-year segments of these residuals, as well as on the timeseries as a whole,
with the goal of finding frequencies that, in conjunction with demographic data,
can predict which districts will change and by how much. I would also like to model
the changes in state congresses with an ARMA model with poisson error and see
if the data support the assumption of stationarity.